---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

*Load the necessary libraries*
```{r}
library("ggplot2")
library("readr")
```


## Loading and preprocessing the data


*We download the file in the working directory and keep only the rows without NAs*
```{r}
Data_full <- read_csv("activity.csv",guess_max = 40000)
Data <- Data_full[complete.cases(Data_full),]
```



## What is mean total number of steps taken per day?

*We create an aggregation from the data frame which calculates the total of the steps for each day then we plot the histogram*
```{r}
m_steps<- aggregate(steps~date,data=Data,sum)
ggplot(data=m_steps, aes(x=steps))+ geom_histogram(bins = 10, color = "white")

paste(" Mean and median are: ",round(mean(m_steps$steps),2)," and ",round(median(m_steps$steps),2))

```


## What is the average daily activity pattern?

*We create an aggregation from the data frame which averages the steps for each interval*
```{r}
m_interval<- aggregate(steps~interval,data=Data,mean)
ggplot(data = m_interval, aes(x = interval, y = steps))+ geom_line(color = "#00AFBB", size = 1)

```


```{r}
paste("The 5-minute interval that, on average, contains the maximum number of steps is : ",m_interval[m_interval$steps == max(m_interval$steps),"interval"]," for a max value of ",max(m_interval$steps))
```



## Imputing missing values

*Here are the numers of NAs for each columns of the full file*
```{r}
colSums(is.na(Data_full))
```
We see that missing values are for steps variable only.


*We will use the mean for the 5 min intervall, that is already stored in the data frame m_interval created above. There a function replaces the values of steps that are NA by the value of m_interval that fits the interval.*
```{r}
Data_full_new <- Data_full

Data_full_new$steps <- mapply(function(x,y) ifelse(is.na(x), m_interval[m_interval$interval==y,"steps"], x), Data_full_new$steps, Data_full_new$interval)

```


*We create an aggregation from the new data frame, which calculates the total of the steps for each day then we plot the new histogram*
```{r}
m_steps2<- aggregate(steps~date,data=Data_full_new,sum)
ggplot(data=m_steps2, aes(x=steps))+ geom_histogram(bins = 10, color = "white")

paste(" Mean and median are: ",round(mean(m_steps2$steps),2)," and ",round(median(m_steps2$steps),2))
```
The histogram is different: it shows that the imputing has added a lot of values above 10000 steps.
The mean stays the same and the median is now eqal to the mean.


## Are there differences in activity patterns between weekdays and weekends?

*We create a new column with the weekday : we first force local in english so that it could be reproductible*
```{r}
Sys.setlocale("LC_ALL","English")
Data_full_new$day <- sapply(Data_full_new$date, function (x) ifelse(weekdays(x)%in%c("Saturday","Sunday"),"weekend","weekday")) 

md_interval<- aggregate(steps~interval+day,data=Data_full_new,mean)
ggplot(data = md_interval, aes(x = interval, y = steps))+ geom_line(color = "#00AFBB", size = 1)+facet_wrap(~day)
```

There is a difference.
